\textmd{\bf Cryptography: Birthday Paradox}

### 2019-01-18 draft.tex
Add outline

    \section{}
    \subsection{}
    Theorem 1.1. Let $S = \{1, 2, \cdots, N\}$. For n times, uniformly randomly draw one element from set S with replacement. Let $x_t$ be the element we draw at time t. Then $\forall p > 0$, there exists a constant $C_1$ such that when $n \geq C_1 \sqrt{N}$, we have
    $$Pr[\exists 1\leq i, j \leq n, i\not = j \text{ such that } x_i = x_j] > p.$$
    \begin{proof}
        Let X denote the event that $\exists i, j \leq t, i\not = j \text{ such that } x_i = x_j$, then we have
        \begin{equation}
            \begin{aligned}
                Pr[\bar{X}] &= \frac{N(N-1)\cdots (N-n+1)}{N^n}\\
                &= \prod_{i=1}^{n-1}(1 - \frac{i}{N})\\
                &\leq \prod_{i=1}^{n-1} exp(-\frac{i}{N})\\
                &= exp(-\frac{n(n-1)}{2N}).
            \end{aligned}
        \end{equation}
        Let $C_1 = \sqrt{-2ln(1 - p)} + 1$. Then when $n \geq C_1\sqrt{N}$, we have
        \begin{equation}
            \begin{aligned}
                n(n-1) > (1 + \sqrt{-2ln(1 - p)\cdot N})(\sqrt{-2ln(1 - p)\cdot N}) > -2ln(1-p)\cdot N.
            \end{aligned}
        \end{equation}
        Which is equivalent to $-\frac{n(n-1)}{2N} < ln(1 - p)$.
        Thus use (1) we have
        $$Pr[\bar{X}] \leq exp(-\frac{n(n-1)}{2N}) < 1 - p.$$
        So we have
        $$Pr[X] > p$$
    \end{proof}
    \subsection{}
    Lemma 1.1. For positive integer $n < N$, we have
    $$\sum_{i=1}^{n-1}ln(1 - \frac{i}{N}) >-\frac{n^2}{N}.$$
    \begin{proof}
        Notice that $\forall x\in [1 -\frac{i+1}{N}, 1 - \frac{i}{N}]$ $(0\leq i\leq n)$, we have $ln(x) < ln(1 - \frac{i}{N})$. Thus
        $$\frac{1}{N}ln(1 - \frac{i}{N}) \geq \int_{1 - \frac{i+1}{N}}^{1 - \frac{i}{N}}ln(x)dx.$$
        Thus we have
        \begin{equation}
            \begin{aligned}
                \frac{1}{N}\sum_{i=1}^{n-1}ln(1 - \frac{i}{N}) &\geq \int_{1-\frac{n}{N}}^1 ln(x)dx\\
                &= (xln(x) - x)|_{1-\frac{n}{N}}^1\\
                &= -\frac{n}{N} - (1 - \frac{n}{N})ln(1 - \frac{n}{N})\\
                &> -\frac{n}{N} - (1 - \frac{n}{N})(-\frac{n}{N})\\
                &= - \frac{n^2}{N^2}.
            \end{aligned}
        \end{equation}
        Thus
        $$\sum_{i=1}^{n-1}ln(1 - \frac{i}{N}) > -\frac{n^2}{N}.$$
    \end{proof}
    Theorem 1.2. Let $S = \{1, 2, \cdots, N\}$. For n times, uniformly randomly draw one element from set S with replacement. Let $x_t$ be the element we draw at time t. Then $\forall p > 0$, there exists a constant $C_2$ such that when $n \leq C_2 \sqrt{N}$, we have
    $$Pr[\exists 1\leq i, j \leq n, i\not = j \text{ such that } x_i = x_j] < p.$$
    \begin{proof}
        Let X denote the event that $\exists i, j \leq t, i\not = j \text{ such that } x_i = x_j$.

>>>>

    \section{Introduction and Outline}
    In this project, we study the Birthday Paradox, in which we uniformly randomly draw n elements with replacement from a set of N elements. The primary goal is to prove a bound of n to promise a collision (some element is drawn for at least 2 times) with some constant probability p. The result is n should be at least $\theta(N^{\frac{1}{2}})$ and we proved this bound cannot be improved. We study this in the first part.

    In the second part, we prove that if the probability of drawing each one of the N elements is not uniformly $\frac{1}{N}$, the probability that a collision occurs is greater or equal to the probability under the uniform case. This provide a plausible reason on why we should use hash functions that have uniform probability and how we can attack them.

    Then we study the bound of n to promise a d-time collision (some element is drawn for at least d times) with some probability p. The result is n should be at least $\theta(N^{\frac{d-1}{d}})$ and we proved this bound cannot be improved.
    \section{}
    \subsection{}
    Theorem 1.1. Let $S = \{1, 2, \cdots, N\}$. For n times, uniformly randomly draw one element from set S with replacement. Let $x_t$ be the element we draw at time t. Then $\forall p > 0$, there exists a constant $C_1$ such that when $n \geq C_1 \sqrt{N}$, we have
    $$Pr[\exists 1\leq i, j \leq n, i\not = j \text{ such that } x_i = x_j] > p.$$
    \begin{proof}
        Let X denote the event that $\exists i, j \leq t, i\not = j \text{ such that } x_i = x_j$, then we have
        \begin{equation}
            \begin{aligned}
                Pr[\bar{X}] &= \frac{N(N-1)\cdots (N-n+1)}{N^n}\\
                &= \prod_{i=1}^{n-1}(1 - \frac{i}{N})\\
                &\leq \prod_{i=1}^{n-1} exp(-\frac{i}{N})\\
                &= exp(-\frac{n(n-1)}{2N}).
            \end{aligned}
        \end{equation}
        Let $C_1 = \sqrt{-2ln(1 - p)} + 1$. Then when $n \geq C_1\sqrt{N}$, we have
        \begin{equation}
            \begin{aligned}
                n(n-1) > (1 + \sqrt{-2ln(1 - p)\cdot N})(\sqrt{-2ln(1 - p)\cdot N}) > -2ln(1-p)\cdot N.
            \end{aligned}
        \end{equation}
        Which is equivalent to $-\frac{n(n-1)}{2N} < ln(1 - p)$.
        Thus use (1) we have
        $$Pr[\bar{X}] \leq exp(-\frac{n(n-1)}{2N}) < 1 - p.$$
        So we have
        $$Pr[X] > p$$
    \end{proof}
    \subsection{}
    Lemma 1.1. For positive integer $n < N$, we have
    $$\sum_{i=1}^{n-1}ln(1 - \frac{i}{N}) >-\frac{n^2}{N}.$$
    \begin{proof}
        Notice that $\forall x\in [1 -\frac{i+1}{N}, 1 - \frac{i}{N}]$ $(0\leq i\leq n)$, we have $ln(x) < ln(1 - \frac{i}{N})$. Thus
        $$\frac{1}{N}ln(1 - \frac{i}{N}) \geq \int_{1 - \frac{i+1}{N}}^{1 - \frac{i}{N}}ln(x)dx.$$
        Thus we have
        \begin{equation}
            \begin{aligned}
                \frac{1}{N}\sum_{i=1}^{n-1}ln(1 - \frac{i}{N}) &\geq \int_{1-\frac{n}{N}}^1 ln(x)dx\\
                &= (xln(x) - x)|_{1-\frac{n}{N}}^1\\
                &= -\frac{n}{N} - (1 - \frac{n}{N})ln(1 - \frac{n}{N})\\
                &> -\frac{n}{N} - (1 - \frac{n}{N})(-\frac{n}{N})\\
                &= - \frac{n^2}{N^2}.
            \end{aligned}
        \end{equation}
        Thus
        $$\sum_{i=1}^{n-1}ln(1 - \frac{i}{N}) > -\frac{n^2}{N}.$$
    \end{proof}
    Theorem 1.2. Let $S = \{1, 2, \cdots, N\}$. For n times, uniformly randomly draw one element from set S with replacement. Let $x_t$ be the element we draw at time t. Then $\forall p > 0$, there exists a constant $C_2$ such that when $n \leq C_2 \sqrt{N}$, we have
    $$Pr[\exists 1\leq i, j \leq n, i\not = j \text{ such that } x_i = x_j] < p.$$
    \begin{proof}
        Let X denote the event that $\exists i, j \leq t, i\not = j \text{ such that } x_i = x_j$.
